¥documentclass[11pt]{article}
¥usepackage[a4paper,margin=1in]{geometry}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage{mathtools}
¥usepackage{bm}
¥usepackage{microtype}
¥usepackage{hyperref}
¥usepackage{xcolor}
¥usepackage{graphicx}

¥title{¥vspace{-1em}¥textbf{The $¥boldsymbol{¥varphi}$-Clock Universe}}
¥author{VR ¥& AI}
¥date{}

¥newtheorem{definition}{Definition}
¥newtheorem{lemma}{Lemma}
¥newtheorem{proposition}{Proposition}
¥newtheorem{theorem}{Theorem}
¥theoremstyle{remark}
¥newtheorem{remark}{Remark}

¥newcommand{¥bb}[1]{¥boxed{¥;#1¥;}}

¥begin{document}
¥maketitle

¥begin{abstract}
We give a complete first-principles formalism for a temporal universe driven by a single bi-infinite binary register, a single tick at $k{=}0$, and the unique value-preserving local rewrite $11¥!¥to¥!100$.
All derivations use only the golden identity $¥varphi^2=¥varphi+1$ (equivalently $1+¥varphi^{-1}=¥varphi$) and this rewrite.
Negative indices are read through a purely temporal inversion lens into outer scale-channel labels $(R{=}¥varphi^m,r{=}(2m)¥!¥!¥pmod3)$, yielding conserved fast content $Q^-=¥sum_{m¥ge1}U_{-m}¥varphi^{-m}$.
Detuned unity admits a $¥varphi$-log spiral envelope; the per-bit one-turn spiral-action is linear in $¥varphi^{-m}$, so the total action is a constant multiple of $Q^-$.
We define species (fast masks), a minimal read-out process into a uniform potential (instantiation), the expected instantiation rate $¥lambda_S=¥sum S(m)U_{-m}¥varphi^{-2m}$, a branching readout that yields exponential growth (``big bang'') without changing the register, and an algebraic binding vs.¥ scatter rule via $¥varphi$-closure (worked example).
A Z$_3$ spinor-like internal cycle arises from residues $r=(2m)¥!¥!¥pmod3$ on depth triplets $¥{m_0,m_0{+}2,m_0{+}4¥}$, with phase advances gated by rung statistics.
We close with a minimal field picture and a simple complexity bound for the tick.
¥end{abstract}

¥section{Axioms and admissible class}¥label{sec:axioms}
Let $¥varphi=¥frac{1+¥sqrt5}{2}$ so $¥varphi^2=¥varphi+1$ and $1+¥varphi^{-1}=¥varphi$.

¥begin{definition}[Register and admissibility]
A register is a bi-infinite binary string
¥[
U=(¥ldots,U_{-1}¥ ¥mid¥ U_0,U_1,U_2,¥ldots),¥qquad U_k¥in¥{0,1¥},
¥]
with the single constraint $U_kU_{k+1}¥neq 11$ (¥emph{no adjacent ones}).
¥end{definition}

¥begin{definition}[Tick and normalization]¥label{def:tick}
The unique value-preserving local rewrite is
¥begin{equation}¥label{rw}
¥underline{11}¥ ¥longrightarrow¥ ¥underline{100}¥qquad¥text{at indices }(k{+}1,k)¥mapsto(k{+}2,k{+}1,k),
¥end{equation}
which restates $¥varphi^{k+1}{+}¥varphi^k=¥varphi^{k+2}$.
A tick injects $1$ at $k=0$ and normalizes:
¥begin{equation}¥label{tick}
U¥ ¥mapsto¥ ¥mathsf{Norm}¥big(U+¥mathbf e_0¥big),¥qquad (¥mathbf e_0)_0=1, ¥ (¥mathbf e_0)_k=0¥ (k¥neq0).
¥end{equation}
¥end{definition}

¥begin{remark}[Why $11¥!¥to¥!100$ and why it halts]
Let $E_q(U):=¥sum_k q^{¥,k}U_k$ for any $q¥in(1,¥varphi)$. Then under ¥eqref{rw} we have
¥[
E_q(100)-E_q(11)=q^{k+2}-(q^{k+1}{+}q^k)=q^k(q^2-q-1)¥ <¥ 0,
¥]
so repeated rewrites strictly decrease $E_q$, hence normalization halts.
¥end{remark}

We work in a natural class that ensures well-defined functionals on the fast tail.

¥begin{definition}[Admissible class]¥label{def:class}
¥(
¥mathcal{C}
:=
¥big¥{U¥ ¥text{admissible}¥ :¥
¥#¥{k¥ge0:U_k=1¥}<¥infty,¥ ¥ ¥sum_{m¥ge1}U_{-m}¥varphi^{-m}<¥infty¥big¥}.
¥)
¥end{definition}

¥section{Uniqueness and locality}

¥paragraph{Normalization order (determinism).}
To resolve the $111$ critical overlap at indices $(-1,0,1)$ we ¥emph{fix} normalization to apply $11¥!¥to¥!100$ from ¥emph{higher to lower} indices (rightmost-first scan).
This makes the post-tick normal form independent of rule order.
With this choice, a boundary rewrite across $(-1,0)$ can occur at the first tick only if $U_{-1}=1$ and $U_{1}=0$ just before the tick.

¥begin{theorem}[Normal form exists and is unique]¥label{thm:nf}
Each tick (insertion at $k=0$) followed by rightmost-first normalization halts and yields a unique normal form.
¥end{theorem}
¥begin{proof}[Sketch]
Halting: the potential $E_q(U)=¥sum_k U_k q^k$ with $1<q<¥varphi$ strictly decreases on $11¥!¥to¥!100$.
Local confluence: the only critical pair is $111$, which our rightmost-first order resolves deterministically by firing $(0,1)$ before $(-1,0)$.
By Newman's lemma (termination $+$ local confluence), normalization is confluent; thus the normal form is unique.
¥end{proof}

¥begin{proposition}[Tick locality]¥label{prop:local-fixed}
Under a tick at $k=0$ and subsequent normalization, digits with $k¥le -2$ never change.
¥end{proposition}
¥begin{proof}
Each rewrite replaces $(k,k{+}1)$ by $(k{+}2,k{+}1,k)$, so the minimum index touched is non-decreasing.
Starting at $k=0$, no chain of carries can cross below $-1$.
¥end{proof}

¥paragraph{Tickability.}
We restrict ticks to states with $U_0=0$ (the slow prefix begins with a $0$); this avoids intermediate digits ``2'' during insertion.


¥begin{lemma}[Uniqueness of normal form (Zeckendorf-like)]¥label{lem:uniq}
If $U,V$ are admissible and $¥sum_k U_k¥varphi^k=¥sum_k V_k¥varphi^k$, then $U=V$.
¥end{lemma}

¥begin{remark}[Boundary probability: ensemble vs.¥ time]¥label{rem:boundary-prob}
A boundary event at the first tick occurs iff $U_{-1}=1$ (and, with the rightmost-first rule, $U_1=0$).
Under the Parry (max-entropy) measure on the golden-mean subshift, the marginal $¥Pr[U_{-1}=1]=¥pi_1=¥frac{1}{1+¥varphi^2}$.
Along any single trajectory the boundary can occur at most once, so its time-average frequency tends to zero; $¥pi_1$ is an ¥emph{ensemble} probability over initial states.
Under a stationary ergodic measure, the boundary event is a zero窶素requency tail event along a single trajectory; probabilities are ensemble quantities.
¥end{remark}


¥section{Projection lens and conserved fast amount}¥label{sec:proj}
We do ¥emph{not} introduce space. The lens is a temporal relabeling of negative bits.
For each $m¥ge1$, if $U_{-m}=1$ we define an outer feature
¥begin{equation}¥label{proj}
(R_m,r_m):=¥big(¥varphi^{m},¥ (2m)¥bmod 3¥big).
¥end{equation}
Define the fast amount
¥begin{equation}¥label{Qminus}
¥bb{¥,Q^-(U):=¥sum_{m¥ge1} U_{-m}¥,¥varphi^{-m}¥,}=¥sum_{(R_m,r_m)}¥frac{1}{R_m}.
¥end{equation}

¥begin{lemma}[Boundary and conservation]¥label{lem:boundary}
Any rewrite $11¥!¥to¥!100$ with $k¥le -2$ leaves $Q^-$ unchanged.
A rewrite at $k=-1$ reduces $Q^-$ by exactly $¥varphi^{-1}$.
¥end{lemma}

¥begin{remark}[Rarity of boundary events]
In a single trajectory the boundary event can happen at most once (it requires $U_{-1}=1$ just before the tick; afterwards $U_{-1}¥leftarrow 0$).
Under the Parry (max-entropy) measure on the golden-mean subshift, $¥Pr[U_{-1}=1]=¥frac{1}{1+¥varphi^2}$; along a single trajectory its time-average frequency tends to zero.
¥end{remark}

¥section{Detuned unity as a ¥texorpdfstring{$¥varphi$}{phi}-log temporal spiral}¥label{sec:spiral}
Let $r(¥theta)=r_¥star e^{b¥theta}$ be a logarithmic spiral.
Its arc length between $¥theta_0$ and $¥theta_1$ is
¥begin{equation}¥label{spiral-length}
L=¥int_{¥theta_0}^{¥theta_1}¥!¥!¥sqrt{r^2+¥Big(¥frac{dr}{d¥theta}¥Big)^2}¥,d¥theta
=¥sqrt{1+b^2}¥int r(¥theta)¥,d¥theta
=¥frac{¥sqrt{1+b^2}}{b}¥,r(¥theta_0)¥big(e^{b(¥theta_1-¥theta_0)}-1¥big).
¥end{equation}
Choose growth per quarter turn $g=e^{b¥pi/2}$ so over one full turn $e^{b(¥theta_1-¥theta_0)}=g^4$.
Associate a fast bit at $-m$ with starting diameter $D_m=¥varphi^{-m}$ (so $r(¥theta_0)=¥tfrac12¥varphi^{-m}$) and orientation $¥sigma¥in¥{+1,-1¥}$ with $g=¥varphi^¥sigma$.
Then the ¥emph{per-bit} one-turn spiral-action is
¥begin{equation}¥label{bit-action}
L_m^{(¥sigma)}=¥frac12¥,¥frac{¥sqrt{1+b_¥sigma^2}}{b_¥sigma}¥,¥big(¥varphi^{4¥sigma}-1¥big)¥ ¥varphi^{-m}
=:C_¥sigma¥,¥varphi^{-m},¥qquad b_¥sigma=¥tfrac{2}{¥pi}¥ln(¥varphi^¥sigma).
¥end{equation}
Summing over the fast tail,
¥begin{equation}¥label{total-action}
¥bb{¥,L_{¥text{turn}}^{(¥sigma)}(U)=¥sum_{m¥ge1}U_{-m}¥,L_m^{(¥sigma)}=C_¥sigma¥,Q^-(U)¥,},
¥end{equation}
a constant multiple of $Q^-$ (changes only at the single boundary by $¥Delta L^{(¥sigma)}=-C_¥sigma¥varphi^{-1}$).

¥section{Slow rungs and scaling}¥label{sec:rungs}
A tick reaches slow rung $n$ iff the pre-tick slow prefix equals the alternating word $(01)^n$.
Under a natural stationary measure on admissible slow prefixes (the golden-mean subshift), the probability of $(01)^n$ is $¥propto¥varphi^{-2n}$, yielding
¥begin{equation}¥label{rung-scaling}
¥bb{¥,f_n¥ ¥propto¥ ¥varphi^{-2n}¥,f_0,¥qquad T_n¥ ¥propto¥ ¥varphi^{2n}¥,T_0¥,}.
¥end{equation}
¥begin{remark}[Measure remark]
The proportionality is asymptotic under any stationary, ergodic measure compatible with no-adjacency; for a fixed initial $U$ it gives a robust envelope for long-run frequencies.
¥end{remark}


¥subsection*{Markov construction for rung statistics}
Let $¥mathsf G$ be the golden-mean subshift forbidding $11$.
The $2¥times 2$ adjacency is $A=¥begin{psmallmatrix}1&1¥¥1&0¥end{psmallmatrix}$ with Perron root $¥lambda=¥varphi$ and right/left eigenvectors $v=(¥varphi,1)$, $w=(1,¥varphi^{-1})$.

The Parry transition matrix is
$P=¥begin{psmallmatrix}¥varphi^{-1} & ¥varphi^{-2}¥¥[2pt] 1 & 0¥end{psmallmatrix}$.
Its stationary distribution is
¥[
¥pi=¥Big(¥frac{¥varphi^2}{1+¥varphi^2},¥ ¥frac{1}{1+¥varphi^2}¥Big),
¥]
so for the alternating prefix $(01)^n$ we have
¥[
¥Pr[(01)^n]=¥pi_0¥,(P_{01}P_{10})^{n}
=¥frac{¥varphi^2}{1+¥varphi^2}¥,¥varphi^{-2n}¥propto ¥varphi^{-2n}.
¥]
Hence the expected waiting time to reach rung $n$ scales as $¥varphi^{2n}$.

¥section{Species schema (genotype)}¥label{sec:species}
A ¥emph{species} is a fast-depth/channel mask: choose $¥mathcal R_S¥subseteq¥{0,1,2¥}$ and $¥mathcal M_S¥subseteq¥mathbb N_{¥ge1}$; set
¥(
S(m)=¥mathbf 1¥{¥,m¥in¥mathcal M_S¥ ¥&¥ (2m)¥bmod3¥in¥mathcal R_S¥,¥}.
¥)
Define observables
¥begin{align}
¥text{Multiplicity:}&¥quad N_S(U)=¥sum_{m¥ge1} S(m)¥,U_{-m},¥¥
¥text{Weight:}&¥quad W_S(U)=¥sum_{m¥ge1} S(m)¥,U_{-m}¥,¥varphi^{-m},¥¥
¥text{Channels:}&¥quad ¥mathbf C_S(U)=(C_0,C_1,C_2),¥ C_j=¥sum_{m¥ge1} S(m)¥,U_{-m}¥,¥mathbf 1¥{(2m)¥bmod3=j¥},¥¥
¥text{Central depth/time-scale:}&¥quad ¥bar m_S=¥frac{¥sum m¥,S(m)¥,U_{-m}¥,¥varphi^{-m}}{¥sum S(m)¥,U_{-m}¥,¥varphi^{-m}},¥quad
T_S¥propto ¥varphi^{2¥bar m_S}T_0,¥¥
¥text{Spiral-action:}&¥quad L_S^{(¥sigma)}=C_¥sigma¥,W_S(U).
¥end{align}
By Proposition‾¥ref{prop:local}, these are constants of motion except for the one-off boundary at $m{=}1$ (if present in $S$).

¥section{Instantiation in a uniform potential (phenotype)}¥label{sec:inst}
The outside ¥emph{foam} is a uniform set of temporal slots $¥mathcal F=¥{(R>1,r¥in¥{0,1,2¥})¥}$.
A tick at time $¥tau$ reaches a random rung $n_¥tau$; it ¥emph{instantiates} all fast bits at depths $m¥le n_¥tau$:
¥begin{equation}
¥mathcal I_¥tau=¥{¥, (R_m=¥varphi^m,¥ r_m) ¥ :¥ U_{-m}=1,¥ m¥le n_¥tau¥,¥}.
¥end{equation}
Under the stationary measure, $¥Pr[n_¥tau¥ge m]=¥varphi^{-2m}$, hence the expected per-tick instantiation count for species $S$ is
¥begin{equation}¥label{lambdaS}
¥bb{¥,¥lambda_S=¥sum_{m¥ge1} S(m)¥,U_{-m}¥,¥varphi^{-2m}¥,},¥qquad ¥mathbb E¥,N_S(T)=T¥,¥lambda_S.
¥end{equation}
¥paragraph{Branching readout (exponential phase, no change to $U$).}
If each instantiation becomes a new reader with probability $¥alpha$ and each reader produces on average $¥lambda_¥ast=¥sum_{m¥ge1}U_{-m}¥varphi^{-2m}$ instantiations per tick, then the expected reader count obeys
¥begin{equation}¥label{branching}
¥mathbb E¥,¥mathcal R_{t+1}=(1+¥alpha¥,¥lambda_¥ast)¥,¥mathbb E¥,¥mathcal R_t¥quad¥Rightarrow¥quad
¥mathbb E¥,¥mathcal R_t=¥mathcal R_0(1+¥alpha¥,¥lambda_¥ast)^t,
¥end{equation}
i.e.¥ an exponential ``big bang'' phase arising purely from readout replication.

¥section{Co-instantiation and binding via ¥texorpdfstring{$¥varphi$}{phi}-closure}

¥paragraph{Greedy $¥varphi$-closure and uniqueness.}
Given a finite multiset of depths $M¥subset¥mathbb N$ with total weight $W=¥sum_{m¥in M}¥varphi^{-m}$, let $m_{¥rm eff}=¥max¥{m:¥varphi^{-m}¥le W¥}$, $N=¥lfloor W¥varphi^{m_{¥rm eff}}¥rfloor$, and $¥gamma=W-N¥varphi^{-m_{¥rm eff}}¥in[0,¥varphi^{-m_{¥rm eff}})$.
Then
¥[
W=N¥varphi^{-m_{¥rm eff}}+¥gamma,
¥]
is unique: shifting $m_{¥rm eff}$ up would force $¥gamma<0$; down would make $¥gamma¥ge ¥varphi^{-m_{¥rm eff}}$.
We bind $N$ composites at depth $m_{¥rm eff}$ and route $¥gamma$ to a massless cascade.

¥label{sec:bind}
The expected ¥emph{co-instantiation} rate of $S_1$ and $S_2$ is
¥begin{equation}¥label{co}
¥bb{¥,¥Lambda_{12}=¥sum_{m¥ge1} S_1(m)¥,S_2(m)¥,U_{-m}¥,¥varphi^{-2m}¥,}.
¥end{equation}
Upon co-instantiation at the same depth $m$ (with channels fixed by $r=(2m)¥bmod3$), we adopt a minimal algebraic outcome:

¥textbf{Interaction Rule (IR).} Given two co-instantiated depths, perform the greedy
$¥varphi$-closure on their total weight. If the resulting multiplicity $N¥ge 1$,
declare binding with multiplicity $N$ at depth $m_{¥mathrm{eff}}$ and record the
remainder $¥gamma$ in the massless channel. If $N=0$, declare scatter.

¥paragraph{Worked example.}
Depths $m$ and $m{+}2$ (non-adjacent in $U$) have weights $¥varphi^{-m}$ and $¥varphi^{-(m+2)}$:
¥begin{equation}¥label{closure}
¥varphi^{-m}+¥varphi^{-(m+2)}=(¥varphi+¥varphi^{-1})¥,¥varphi^{-(m+1)}=¥sqrt5¥,¥varphi^{-(m+1)}.
¥end{equation}
We declare a ¥emph{composite} at depth $m{+}1$ (channel rotates by $+2$ mod 3) plus a radiative remainder of $(¥sqrt5-1)¥varphi^{-(m+1)}$ in $¥gamma$:
¥[
A_m+B_{m+2}¥ ¥longrightarrow¥ C_{m+1}+¥gamma.
¥]
Action bookkeeping: $L^{(+)}(A_m)+L^{(+)}(B_{m+2})=L^{(+)}(C_{m+1})+L^{(+)}(¥gamma)$ by linearity in ¥eqref{bit-action}.

¥section{¥texorpdfstring{$¥mathrm Z_3$}{Z3} spinor-like cycle (explicit construction)}¥label{sec:spinor}
Let a spinor species use depths $¥mathcal M_{¥mathrm{sp}}=¥{m_0,m_0{+}2,m_0{+}4¥}$ (respecting no-adjacency).
Channels cycle:
¥(
r(m_0{+}2k)=(2m_0{+}4k)¥bmod3=r(m_0)+k¥ (¥bmod3).
¥)
Define an internal phase $p¥in¥{0,1,2¥}$ as the index of the most recently activated component; when a tick reaches the next component (depth increases by $2$), the phase advances
¥begin{equation}
p¥ ¥mapsto¥ p+1¥quad(¥bmod 3).
¥end{equation}
The probability per tick to activate depth $m_0{+}2k$ is $¥varphi^{-2(m_0+2k)}$, hence the expected waiting time for each phase step is
¥begin{equation}
¥mathbb E[¥Delta t_k]=¥varphi^{2(m_0+2k)}.
¥end{equation}
This gives a Z$_3$ internal cycle with geometrically increasing periods---spinor-like and purely temporal (not SU(2)).

¥section{Minimal field picture (optional bridge)}¥label{sec:field}
Let $¥psi_r(m;t)$ be the accumulated instantiations in the foam.
The tick defines a source
¥begin{equation}
J_r(m;t)=U_{-m}¥,¥mathbf 1¥{n_t¥ge m¥}¥,¥mathbf 1¥{r=(2m)¥bmod3¥},
¥end{equation}
and a free update $¥psi_r(m;t{+}1)=¥psi_r(m;t)+J_r(m;t)$.
A minimal interaction (matching ¥eqref{closure}) is a $¥varphi$-closure vertex
¥begin{equation}
¥psi_{r+2}(m{+}1;t{+}1)¥ ¥leftarrow¥ ¥kappa¥ ¥psi_{r}(m;t)¥,¥psi_{r+1}(m{+}2;t),
¥end{equation}
with a massless channel $¥gamma$ recording the radiative remainder.
No spatial geometry is assumed; this is a discrete, scale-indexed skeleton compatible with field-theoretic language.



¥subsection*{Operational distance and synchronization}
Define the shared-prefix length $¥ell(U,V)=¥max¥{L:¥ U_{-m}=V_{-m}¥ ¥forall m¥le L¥}$ and the ultrametric
¥[
d_¥varphi(U,V)=¥varphi^{-¥ell(U,V)}.
¥]
An operational estimator from ticks is obtained by the shortest time $T^¥ast$ needed for both registers to exhibit the same rung $n$; by $¥mathbb E[T_n]¥propto¥varphi^{2n}$ we set
¥(
¥widehat d_¥varphi¥sim (T^¥ast/T_0)^{-1/2}.
¥)

¥section{Algorithmic complexity}

¥subsection*{Deterministic tick harness (for simulation)}
¥begin{enumerate}
¥item Require $U_0=0$; set $U_0¥leftarrow 1$ (tick).
¥item For $k$ descending from the largest index with $U_k=1$ down to $-1$, apply $11¥!¥to¥!100$ at $(k{-}1,k)$ whenever present.
¥item Record: whether $(-1,0)$ fired (boundary), the rung $n$ (length of initial alternating $01$), and any species readouts.
¥end{enumerate}
This runs in $O(n)$ when the slow prefix is $(01)^n$.


Let the pre-tick slow prefix length be $n$ (maximal $(01)^n$). Normalization performs $O(n)$ rewrites and touches indices $¥{0,1,¥dots,O(n)¥}$.
Negative indices $k¥le -2$ are never touched; the boundary $-1$ can be cleared once.
Hence a tick has worst-case time $O(n)$ and space $O(n)$ on the slow side.

¥section{One-page recipe}
¥begin{itemize}
¥item ¥textbf{Engine:} single tick at $k=0$, single rewrite $11¥to100$.
¥item ¥textbf{Lens:} fast bit $-m$ $¥mapsto$ $(R=¥varphi^m,¥ r=(2m)¥bmod3)$.
¥item ¥textbf{Conserved:} $Q^-=¥sum U_{-m}¥varphi^{-m}$ (changes only by $-¥varphi^{-1}$ at $(-1,0)$).
¥item ¥textbf{Spiral-action:} $L^{(¥sigma)}=C_¥sigma Q^-$ with $C_¥sigma$ from ¥eqref{bit-action}.
¥item ¥textbf{Rungs:} $f_n¥propto¥varphi^{-2n}$, $T_n¥propto¥varphi^{2n}$ (natural measures).
¥item ¥textbf{Species:} $(¥mathcal R_S,¥mathcal M_S)¥Rightarrow N_S,W_S,¥mathbf C_S,¥bar m_S, L_S$ (constants of motion).
¥item ¥textbf{Instantiation:} per-tick rate $¥lambda_S=¥sum S(m)U_{-m}¥varphi^{-2m}$; with branching readers ¥eqref{branching} gives exponential growth.
¥item ¥textbf{Interactions:} co-rate $¥Lambda_{12}=¥sum S_1S_2¥,U_{-m}¥varphi^{-2m}$; bind iff $¥varphi$-closure to a single depth exists; else scatter.
¥item ¥textbf{Spinor:} depths $m_0,m_0{+}2,m_0{+}4$ give a Z$_3$ phase advancing when rung reaches the next depth.
¥end{itemize}

¥section*{Figure}
¥begin{figure}[h!]
  ¥centering
  ¥includegraphics[width=¥linewidth]{phi_clock_universe_species_figure.pdf}
  ¥caption{Tick at $k=0$ and the unique normalization $11¥!¥to¥!100$ across $(-1,0)$ remove the nearest fast bit and promote one unit to the slow side, changing $Q^-$ by $-¥varphi^{-1}$.
  The inversion lens re-labels a fast bit at $-m$ as $(R=¥varphi^{m},¥,r=(2m)¥bmod3)$.
  The view is purely temporal (no spatial postulate).}
¥end{figure}

¥section*{Consistency and minimalism}
Every equation is a consequence of $¥varphi^2=¥varphi+1$ and $11¥to100$, plus standard calculus for the log-spiral length ¥eqref{spiral-length} (choice of units enters only through starting diameter).
No extra forces, spaces, or Hamiltonians are assumed; if a spatial metric is ever desired, it must be constructed operationally from synchronization times, not postulated.


% ---------- Appendix: Concrete Examples (input this file) ----------
¥appendix
¥section*{Appendix: Concrete Examples}

¥subsection*{A1. A small register and its invariants}
Consider the admissible register with fast bits at depths $m=1$ and $m=3$ (i.e.¥ $U_{-1}=U_{-3}=1$) and no slow bits. Then
¥begin{align*}
Q^-(U) &= ¥varphi^{-1}+¥varphi^{-3} ¥;=¥; 0.618034+0.236068 ¥;=¥; 0.854102,¥¥
L_{¥text{turn}}^{(+)}(U) &= C_+¥,Q^- ¥quad¥text{with } C_+=¥tfrac12¥frac{¥sqrt{1+b_+^2}}{b_+}(¥varphi^4-1)
¥approx 9.992927 ¥;¥Rightarrow¥; L_{¥text{turn}}^{(+)}¥approx 8.534978,¥¥
L_{¥text{turn}}^{(-)}(U) &= C_-¥,Q^- ¥quad¥text{with } C_-=¥tfrac12¥frac{¥sqrt{1+b_-^2}}{b_-}(¥varphi^{-4}-1)
¥approx 1.457948 ¥;¥Rightarrow¥; L_{¥text{turn}}^{(-)}¥approx 1.245237.
¥end{align*}
(Here $b_¥pm=¥tfrac{2}{¥pi}¥ln(¥varphi^¥pm)$.) These numbers match Eq.¥,( ¥ref{total-action} ): action is linear in $Q^-(U)$.

¥subsection*{A2. One tick and the slow rung}
Let the pre-tick slow prefix be $(01)^2$, i.e.¥ $(U_0,U_1,U_2,U_3)=(0,1,0,1)$. Injecting a tick at $k=0$ and applying only $11¥!¥to¥!100$:
¥begin{align*}
¥cdots¥mid ¥underline{0,1},0,1 ¥;+¥; ¥mathbf e_0 &¥;¥Rightarrow¥; ¥cdots¥mid ¥underline{1,1},0,1 ¥;¥xRightarrow{11¥to100}¥; ¥cdots¥mid 0,0,¥underline{1,1}
¥¥[-0.25em]&¥;¥xRightarrow{11¥to100}¥; ¥cdots¥mid 0,0,0,0,1,
¥end{align*}
so the tick reaches ¥emph{rung} $n=2$ and leaves a single $1$ at index $2n=4$ (all intermediate slow indices cleared).

¥subsection*{A3. Instantiation rate for a species}
Let the species mask $S$ select depths $¥{1,3¥}$. Then the expected instantiations per tick are
¥begin{equation*}
¥lambda_S=¥sum_{m¥in¥{1,3¥}}¥varphi^{-2m}=¥varphi^{-2}+¥varphi^{-6}
= 0.381966+0.055728= 0.437694.
¥end{equation*}
Over $T$ ticks the expected count is $T¥,¥lambda_S$. The ¥emph{rest content} of this species is
¥begin{equation*}
W_S=¥sum_{m¥in¥{1,3¥}}¥varphi^{-m}=¥varphi^{-1}+¥varphi^{-3}=0.854102,¥qquad
L_S^{(¥sigma)}=C_¥sigma¥,W_S.
¥end{equation*}

¥subsection*{A4. Binding vs.¥ scatter by $¥varphi$-closure (worked)} 
Two co-instantiated depths $m$ and $m¥!+¥!2$ have weights $¥varphi^{-m}$ and $¥varphi^{-(m+2)}$. Their sum is
¥begin{equation*}
¥varphi^{-m}+¥varphi^{-(m+2)}=(¥varphi+¥varphi^{-1})¥varphi^{-(m+1)}=¥sqrt5¥,¥varphi^{-(m+1)}.
¥end{equation*}
Thus we declare a composite at depth $m{+}1$ plus a ¥emph{massless} remainder $¥gamma$ of size $(¥sqrt5-1)¥varphi^{-(m+1)}$.
For $m=2$ one gets numerically
¥begin{align*}
¥varphi^{-m}+¥varphi^{-(m+2)}&=0.527864,¥¥
¥text{composite weight at $m{+}1$}&=¥varphi^{-(m+1)}=0.236068,¥¥
¥gamma¥text{ (radiative remainder)}&=(¥sqrt5-1)¥varphi^{-(m+1)}=0.291796.
¥end{align*}
Action bookkeeping follows from linearity: $L^{(+)}$ of inputs equals $L^{(+)}$ of outputs.

¥subsection*{A5. Z$_3$ spinor cycle (explicit)} 
Choose $m_0=2$, so the triplet depths are $¥{2,4,6¥}$ with residues
¥(
r=¥{1,2,0¥}=¥{(2m)¥bmod3¥}.
¥)
Each activation advances the internal phase $p¥mapsto p+1$ (mod 3). Expected waiting times per phase step are
¥begin{equation*}
¥mathbb E[¥Delta t_k]=¥varphi^{2(m_0+2k)}=¥{6.854102,¥,46.978714,¥,321.996894¥}.
¥end{equation*}

¥subsection*{A6. Branching readout (``big bang'' phase)} 
If every instantiation becomes a new reader with probability $¥alpha=0.120000$ and a reader's mean per-tick output is
¥(
¥lambda_¥ast=¥sum_{m=1}^4¥varphi^{-2m}=0.604878,
¥)
then the expected reader count grows as
¥begin{equation*}
¥mathbb E¥,¥mathcal R_t=¥mathcal R_0¥,(1+¥alpha¥lambda_¥ast)^t
=¥mathcal R_0¥,(1.072585)^t.
¥end{equation*}
This is exponential, while the register and its law remain unchanged.


¥end{document}
